Make any stupid plot you want

### 2021-05-10 draft/sections/math.tex
graphics

\subsection{Graphic Space \gtotal}
\label{sec:math:graphic} 
To establish that the artist is a structure preserving map from data \dtotal\ to graphic \gtotal\, we construct a graphic bundle so that we can define \textit{equivariance} in terms of maps on the fiber spaces and \textit{continuity} in terms of maps on the base space. As with the data, we can represent the target graphic as a section \gsection\ of a bundle  $(\gtotal, \gbase, \pi, \gfiber)$. 
\begin{equation}
    \begin{tikzcd}[ampersand replacement=\&]
        \gfiber \arrow[r, hook] \& \gtotal \arrow[d, "\pi"'] \\
                          \& \gbase \arrow[u, "\gsection"', bend right]
    \end{tikzcd}
  \end{equation}
The graphic bundle \gtotal\ consists of a base \gbase (\autoref{sec:math:graphic:fiber}) that is a thickened form of \dbase\, a fiber \gfiber (\autoref{sec:math:graphic:base}) that is an idealized display space, and sections \gsection (\autoref{sec:math:graphic:section}) that encode a graphic where the visual characteristics are fully specified.

\subsubsection{Idealized Display \gfiber}
\label{sec:math:graphic:fiber}
To fully specify the visual characteristics of the image, we construct a fiber \gfiber\ that is a non-pixelated version of the target space. Typically \gtotal\ is trivial and therefore sections can be thought of as mappings into \gfiber. In this work, we assume a 2D opaque image $\gfiber=\reals^5$ with elements 
\begin{equation*}
(x,\, y,\, r,\, g,\, b) \in \gfiber
\end{equation*}
such that a rendered graphic only consists of 2D position and color. To support overplotting and transparency, the fiber could be $\gfiber=\reals^{7}$ such that $(x, y, z, r, g, b, a) \in \gfiber$ specifies the target display. By abstracting the target display space as \gfiber, the model can support different targets, such as a 2D screen or 3D printer.

@@

For a visualization component to preserve continuity, we propose that there must exist a structure preserving surjective map $\vindex: \gbase\rightarrow\dbase$ from the data base space \dbase\ to the graphic base space \gbase. Formally, we require that \dbase\ be a deformation retract\cite{RetractionTopology2020} of \gbase\ such that \dbase\ and \gbase\ have the same homotopy, meaning there is a continuous map from \gbase\ to \dbase\cite{weissteinHomotopy}. The surjective map $\vindex: \gbase \rightarrow \dbase$ 
\begin{equation}
    \begin{tikzcd}
        \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
        \dbase                   & \gbase \arrow[l, "\vindex"']
    \end{tikzcd}
    \label{eq:math:graphic:vindex}
\end{equation}
goes from region $\gbasepoint \in \gbase_{\dbasepoint}$ to its associated point $\dbasepoint \in \dbase$. This means that if $\vindex(\gbasepoint) = \dbasepoint$, the record at \dbasepoint\ is copied over the region \gbasepoint\ such that $\dsection(\dbasepoint)=\vindex^*\dsection(\gbasepoint)$ where $\vindex^*\dsection(\gbasepoint)$  is \dsection\ pulled back over \gbase. The map \vindex\ is part of the implementation of the artist $\mathcal{\vartist}$ and therefore is not defined in terms of the data; instead it is how we specify the constraint that the type of the graphic \textit{continuity} must be able to map to the type of the data \textit{continuity}.

@@

To encode the continuity of the elements in the display fiber \gfiber, the graphic base space \gbase\ has the same dimensionality as the target output space. For example, in \autoref{fig:math:graphic:retraction:map} the base space \gbase\ is a representation of a region of a 2D display space. Since \gbase\ must have the same dimensionality as the output graphic, it is allowed to add dimensions to \dbase\ to make \dbase\ renderable. A point that is 0D in \dbase\ cannot be represented on screen unless it is thickened to 2D (a) to encode the connectivity of the pixels that visually represent the point. This is also the case with the line (b), which would be infinitely think on screen if \gbase\ was not thickened to 2D. This thickening is often not necessary when the dimensionality of \dbase\ matches the dimensionality of the target space, for example if \dbase\ is 2D and the display is a 2D screen (c). Since the mapping function \vindex\ binds the graphic base space to the data base space, it can be used by interactive visualization components to look up the data associated with a region on screen.  One example is to fill in details in a hover tooltip, another is to convert region selection (such as zooming) on \gbase\ to a query on the data to access the corresponding record components on \dbase.

\subsubsection{Graphic}
\label{sec:math:graphic:section}
The section $\gsection: \gbase \rightarrow \gtotal$ is the graphic in an idealized prerender space and also acts as a specification for rendering the graphic to target display format. To demonstrate the role of \gsection\, it is sufficient to sketch out how an arbitrary pixel would be rendered, where a pixel $p$ in a real display corresponds to a region $\gbase_p$ in the idealized display. To determine the color of the pixel, we aggregate the color values over the region via integration:
\begin{align*}
    \label{eq:math:graphic:section:color}
    r_p &= \iint\limits_{S_p} \rho_r(s)ds^{2}\\
    g_p &= \iint\limits_{S_p} \rho_g(s)ds^{2}\\
    b_p &= \iint\limits_{S_p} \rho_b(s)ds^{2}
\end{align*}
For a 2D screen, the pixel is defined as a region $p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]$ of the rendered graphic. Since the x and y in $p$ are in the same coordinate system as the x and y components of \gfiber\,  the inverse map of the bounding box $\gbase_{p} ={\gsection_{x,y}}^{-1}(p)$ is a region $\gbase_p \subset \gbase$. The color is the result of the integration over  $\gbase_p$.

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/render.png}
    \caption{To render a graphic, a pixel $p$ is selected in the display space, which is defined in the same coordinates as the x and y components in \gfiber\ via the renderer. In \gtotal\, the inverse mapping ${\gsection_{x,y}}^(p)$ returns a region $\gbase_{p} \subset \gbase$. $\gsection(\gbase_{p})$ returns a set of points $(x,\,y,\,r,\,g,\,b) \in \gfiber$ that lie over $\gbase_{p}$. The integral over the $(r,\,g,\,b)$ pixels specifies that the pixel should be green}
    \label{fig:math:graphic:rho:lookup}
\end{figure}

As shown in \autoref{fig:math:graphic:rho:lookup}, a pixel $p$ in the output space, drawn in yellow, is selected and mapped, via the renderer, into a region on \gtotal.  The region on \gtotal\ corresponds to a region $\gbase_p \subset \gbase$ via the inverse mapping  ${\gsection_{xy}}^(p)$. The base space \gbase\ is an annulus to match the topology of the graphic idealized in \gtotal. The section  $\gsection(\gbase_{p})$  then maps into the fiber \gfiber\ over $\gbase_p$ to obtain the set of points in \gfiber, here represented as a table, that correspond to that section. The integral over the pixel components of this set of points in the fiber yields the color of the pixel. In general, \gsection\ is an abstraction of rendering. In very broad strokes \gsection\ can be a specification such as PDF\cite{bienz1993portable}, SVG\cite{quintScalable2003}, or an openGL scene graph\cite{CarsonOpenGL1997} or a rendering engine such as cairo\cite{CairographicsOrg} or AGG\cite{shemanarevAntiGrainGeometry}. Implementation of \gsection\ is out of scope for this proposal.

>>>>

\subsection{Graphic Space \gtotal}
\label{sec:math:graphic} 
To establish that the artist is a structure preserving map from data \dtotal\ to graphic \gtotal\, we construct a graphic bundle so that we can define \textit{equivariance} in terms of maps on the fiber spaces and \textit{continuity} in terms of maps on the base space. As with the data \dsection, we can represent the target graphic as a section \gsection\ of a bundle  \((\gtotal, \gbase, \pi, \gfiber)\). 
\begin{equation}
    \begin{tikzcd}[ampersand replacement=\&]
        \gfiber \arrow[r, hook] \& \gtotal \arrow[d, "\pi"'] \\
                          \& \gbase \arrow[u, "\gsection"', bend right]
    \end{tikzcd}
  \end{equation}
The graphic bundle \gtotal\ consists of a base \gbase (\autoref{sec:math:graphic:fiber}) that is a thickened form of \dbase\, a fiber \gfiber (\autoref{sec:math:graphic:base}) that is an idealized display space, and sections \gsection (\autoref{sec:math:graphic:section}) that encode a graphic where the visual characteristics are fully specified.

\subsubsection{Idealized Display \gfiber}
\label{sec:math:graphic:fiber}
To fully specify the visual characteristics of the image, we construct a fiber \gfiber\ that is a non-pixelated version of the target space. Typically \gtotal\ is trivial and therefore sections can be thought of as mappings into \gfiber. In this work, we assume a 2D opaque image $\gfiber=\reals^5$ with elements 
\begin{equation*}
(x,\, y,\, r,\, g,\, b) \in \gfiber
\end{equation*}
such that a rendered graphic only consists of 2D position and color. To support overplotting and transparency, the fiber could be \(\gfiber=\reals^{7}\) such that \((x, y, z, r, g, b, a) \in \gfiber\) specifies the target display. By abstracting the target display space as \gfiber, the model can support different targets, such as a 2D screen or 3D printer.

@@

For a visualization component to preserve continuity, we propose that there must exist a structure preserving surjective map \(\vindex: \gbase\rightarrow\dbase\) from the data base space \dbase\ to the graphic base space \gbase. Formally, we require that \dbase\ be a deformation retract\cite{RetractionTopology2020} of \gbase\ such that \dbase\ and \gbase\ have the same homotopy, meaning there is a continuous map from \gbase\ to \dbase\cite{weissteinHomotopy}. The surjective map \(\vindex: \gbase \rightarrow \dbase\)
\begin{equation}
    \begin{tikzcd}
        \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
        \dbase                   & \gbase \arrow[l, "\vindex"']
    \end{tikzcd}
    \label{eq:math:graphic:vindex}
\end{equation}
goes from region \(\gbasepoint \in \gbase_{\dbasepoint}\) to its associated point \(\dbasepoint \in \dbase\). This means that if \(\vindex(\gbasepoint) = \dbasepoint\), the record at \dbasepoint\ is copied over the region \gbasepoint\ such that \(\dsection(\dbasepoint)=\vindex^*\dsection(\gbasepoint)\) where \(\vindex^*\dsection(\gbasepoint)\)  is \dsection\ pulled back over \gbase. The map \vindex\ is part of the implementation of the artist \(\mathcal{\vartist}\) and therefore is not defined in terms of the data; instead it is how we specify the constraint that the type of the graphic \textit{continuity} must be able to map to the type of the data \textit{continuity}.

@@

To encode the continuity of the elements in the display fiber \gfiber, the graphic base space \gbase\ has the same dimensionality as the target output space. For example, in \autoref{fig:math:graphic:retraction:map} the base space \gbase\ is a representation of a region of a 2D display space. Since \gbase\ must have the same dimensionality as the output graphic, it is allowed to add dimensions to \dbase\ to make \dbase\ renderable. A point that is 0D in \dbase\ cannot be represented on screen unless it is thickened to 2D (a) to encode the connectivity of the pixels that visually represent the point. This is also the case with the line (b), which would be infinitely thin on screen if \gbase\ was not thickened to 2D. This thickening is often not necessary when the dimensionality of \dbase\ matches the dimensionality of the target space, for example if \dbase\ is 2D and the display is a 2D screen (c). Since the mapping function \vindex\ binds the graphic base space to the data base space, it can be used by interactive visualization components to look up the data associated with a region on screen.  One example is to fill in details in a hover tooltip, another is to convert region selection (such as zooming) on \gbase\ to a query on the data to access the corresponding record components on \dbase.

\subsubsection{Graphic}
\label{sec:math:graphic:section}
The section \(\gsection: \gbase \rightarrow \gtotal\) is the graphic in an idealized prerender space and also acts as a specification for rendering the graphic to target display format. To demonstrate the role of \gsection\, it is sufficient to sketch out how an arbitrary pixel would be rendered, where a pixel \(p\) in a real display corresponds to a region \(\gbase_p\) in the idealized display. To determine the color of the pixel, we aggregate the color values over the region via integration:
\begin{align*}
    \label{eq:math:graphic:section:color}
    r_p &= \iint\limits_{S_p} \rho_r(s)ds^{2}\\
    g_p &= \iint\limits_{S_p} \rho_g(s)ds^{2}\\
    b_p &= \iint\limits_{S_p} \rho_b(s)ds^{2}
\end{align*}
For a 2D screen, the pixel is defined as a region \(p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]\) of the rendered graphic. Since the x and y in \(p\) are in the same coordinate system as the x and y components of \gfiber\,  the inverse map of the bounding box \(\gbase_{p} ={\gsection_{x,y}}^{-1}(p)\) is a region \(\gbase_p \subset \gbase\). The color is the result of the integration over  \(\gbase_p\).

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/render.png}
    \caption{To render a graphic, a pixel \(p\) is selected in the display space, which is defined in the same coordinates as the x and y components in \gfiber\ via the renderer. Therefore, the pixel \(p\) maps to a region on \gtotal. In \gtotal\, the inverse mapping \({\gsection_{x,y}}^(p)\) returns a region \(\gbase_{p} \subset \gbase\). \(\gsection(\gbase_{p})\) returns a set of points \((x,\,y,\,r,\,g,\,b) \in \gfiber\) that lie over \(\gbase_{p}\). The integral over the \((r,\,g,\,b)\) pixels specifies that the pixel should be green}
    \label{fig:math:graphic:rho:lookup}
\end{figure}

As shown in \autoref{fig:math:graphic:rho:lookup}, a pixel \(p\) in the output space, drawn in yellow, is selected and mapped, via the renderer, into a region on \gtotal.  The region on \gtotal\ corresponds to a region \(\gbase_p \subset \gbase\) via the inverse mapping  \({\gsection_{xy}}^(p)\). The base space \gbase\ is an annulus to match the topology of the graphic idealized in \gtotal. The section  \(\gsection(\gbase_{p})\) then maps into the fiber \gfiber\ over \(\gbase_p\)to obtain the set of points in \gfiber, here represented as a table, that correspond to that section. The integral over the pixel components of this set of points in the fiber yields the color of the pixel. In general, \gsection\ is an abstraction of rendering. In very broad strokes \gsection\ can be a specification such as PDF\cite{bienz1993portable}, SVG\cite{quintScalable2003}, or an openGL scene graph\cite{CarsonOpenGL1997} or a rendering engine such as cairo\cite{CairographicsOrg} or AGG\cite{shemanarevAntiGrainGeometry}. Implementation of \gsection\ is out of scope for this proposal.
